\tit{Waarmee} vs. \tit{met wat}

### 2019-12-24 paper.tex
first draft
finished it

\title{Dutch, PPs, ʀ-pronouns}
\author{Fenna Bergsma}
\date{\today}

@@

The term \tsc{r}-pronoun \citep{riemsdijk1978} refers to a set of nominal elements that can strand prepositions in Dutch (and German). It is special because Dutch is a normally non-preposition stranding language. \tsc{R}-pronouns morphologically consist of a combination of a preposition and locative pronoun. In the example \ref{ex:klimerop} \tit{mee} `with' is the preposition and \tit{er} `there' is the locative pronoun.

@@

This paper focuses on a single \tsc{r}-pronoun in a specific type of free relative construction. The \tsc{r}-pronoun I focus on is the relative pronoun \tit{waarmee} `with what', which is interesting for two reasons. First, just like for all \tsc{r}-pronouns, the wh-element is the locative, but there is no meaning component related to location in `with what'. Second, the preposition \tit{met} `with' changes into \tit{mee} `with when it is combined with an \tsc{r}-pronoun.

The construction I focus on is a mismatching free relative. This is a free relative construction in which the two predicates (the one in the main clause and the one in the embedded clause) combine with two different cases (i.e. the case requirements do not match). I illustrate this in \ref{ex:gekochtwaarmee}. The predicate in the embedded clause, \tit{schildert} `paint', combines with an instrumental PP. The predicate in the main clause clause, \tit{gekocht} `bought' combines with an accusative DP. The \tsc{r}-pronoun \tit{waarmee} `with what' is used here.
\footnote{In this example \tit{mee} `with' is stranded and \tit{waar} `where' is moved to the left edge of the embedded clause. It is also possible for \tit{waar} `where' to bring \tit{mee} `with' along, as in \ref{ex:meealong} but it is regarded as slightly less natural.

@@

If one were to switch around the predicates between the clauses, the \tsc{r}-pronoun does not appear anymore. In \ref{ex:schildermet}, \tit{schilder} `paint' combines with an instrumental PP in the main clause and \tit{gekocht} `bought' combines with an accusative DP in the embedded clause. The use of an \tsc{r}-pronoun is ungrammatical, as indicated by the ungrammaticality of \ref{ex:schilderwaarmee}. Instead, a combination of the regular instrumental preposition \tit{met} `with' and the regular wh-pronoun \tit{was} `what' in used.

@@

In this paper I show that distribution of \tit{waarmee} `with what' and \tit{met wat} `with what' in these free relative constructions gives us a unique insight into the internal structure of \tsc{r}-pronouns. In what follows I show that \tsc{r}-pronouns and regular preposition compete to spell out the same syntactic features. If all features form a constituent, the \tsc{r}-pronoun surfaces. If the constituent is interrupted, the preposition-pronoun combination shows up. This straightforwardly follows in a system in which spellout targets phrasal constituents: Nanosyntax \citep{starke2009}.

This paper is structured as follows. First, I show that it really is constituency. Then I decompose \tsc{r}-pronouns, prepositions and regular pronouns. Last, I show in derivations that constituency connects to the choice for \tsc{r}-pronoun or preposition and regular pronoun. Unmarked examples are constructed and have been verified by native speakers.

\section{The distribution between \tit{waarmee} and \tit{met wat}}

In the introduction I discussed the distribution between \tit{waarmee} `with what' and \tit{met wat} `with what' in free relatives with predicates that combine with different cases. Table \ref{tbl:distribution} repeats the generalization. When the main clause predicate combines with an accusative and the embedded clause predicate with an instrumental, \tit{waarmee} is grammatical and \tit{met wat} is ungrammatical. When the main clause predicate combines with an instrumental and the embedded clause predicate with an accusative, \tit{waarmee} is ungrammatical and \tit{met wat} is used.

\begin{table}[ht]\label{tbl:distribution}
	\center
	\caption {Distribution between \tit{waarmee} and \tit{met wat}}
	\begin{minipage}{0.45\linewidth}
		\begin{tabularx}{\textwidth}{ccc}
		\toprule
                              & \tit{waarmee} & \tit{met wat} \\
		\midrule
    m:\tsc{acc}, e:\tsc{ins}  & ✔             & *             \\
    m:\tsc{ins}, e:\tsc{acc}  & *             & ✔             \\
    \bottomrule
\end{tabularx}
\end{minipage}
\end{table}

@@

 \ex. Objects of verbs \label{ex:objverb}
 \ag. Ik zie haar/hem.\\
  I see her/him\\
  `I see her/him.'\label{ex:aniobj}
 \bg. Ik zie 't.\\
  I see it\\
  `I see it.'\label{ex:inaniobj}

@@

\ex. Objects of prepositions \label{ex:objprep}
\ag. Ik schilder samen met haar/hem.\\
 I paint together with her/him\\
 `I am painting together with her/him.'\label{ex:prepani}
\bg. *Ik schilder met 't.\\
 I paint with it\\
 `I am painting with it.'\label{ex:prephet}
\bg. Ik schilder er-mee.\\
 I paint there-with\\
 `I am painting with it.'\label{ex:preper}
\bg. *Ik schilder mee er.\\
 I paint with-there\\
 `I am painting with it.'\label{ex:erprep}

@@

The situation of the inanimate wh-pronouns resembles the inanimate personal pronouns. \tit{Wat} `what' can function as an object of a verb (see \ref{ex:wat}), but not as an object of a preposition \ref{ex:metwat}. In that case, the \tsc{r}-pronoun \tit{waarmee} `with what' is appears, as shown in \ref{ex:waarmee}.
\footnote{The sentence in \ref{ex:metwat} is unacceptable with neutral intonation. It becomes is only acceptable if \tit{wat} `what' is stressed, for example in a context in which the speaker questions the choice for the object hearer is painting with.}

@@

\begin{table}[ht]\label{tbl:inanimates}
	\center
	\caption {Inanimates in Dutch}
	\begin{minipage}{0.7\linewidth}
		\begin{tabularx}{\textwidth}{ccc}
		\toprule
                        & pers. pronouns & wh-pronouns \\
  	\midrule
objects of verbs        & ət             & wat         \\
objects of prepositions & er             & waar        \\
\bottomrule
\end{tabularx}
\end{minipage}
\end{table}

@@

In each of these examples the instrumental PP forms a constituent at a certain point in the derivation. In \ref{ex:const1}, the PP forms a proper constituent in the surface order, as shown in \ref{ex:const1stage}. In \ref{ex:const2}, the PP forms a proper constituent before wh- and V2- movement, shown in \ref{ex:const2stage}.
The structure in \ref{ex:const3stage} represents a stage in the derivation of the embedded clauses in \ref{ex:const3} and \ref{ex:const4}.
\footnote{I assume that the antecedent in a free relative is a phonologically empty element, in line with cf. \cite{bresnan1978a,groos1981,himmelreich2017}. Under that view, \ref{ex:const3} and \ref{ex:const4} are identical, except for that in \ref{ex:const4} modifies phonologically empty element.}
Again, in the stage, which comes before relative movement of the pronoun to the left periphery of the relative clause, the PP forms a proper constituent.

@@

\ex.\label{ex:constwaarmee}
\ag. Waarmee schilder jij?\\
 {where with} paint you\\
 `What are you painting with?'
\bg. Ik schilder [waarmee jij ook schildert].\\
 I paint {where with} you also paint\\
 `I am painting with what you are painting with too.'
\bg. Ik schilder met de kwast [waarmee jij ook schildert].\\
 I paint with the brush {where with} you also paint\\
 `I am painting with the brush that you are painting with too.'

@@

Let me now show how this applies to the examples with the mismatching free relatives. The two predicates I used in the free relatives are \tit{kopen} `to buy' and \tit{schilderen}  `to paint'. \tit{Kopen} `to buy' takes an accusative DP as its object, illustrated in \ref{ex:kopen}. \tit{Schilderen} `to paint' can take an instrumental PP as its object., shown in \ref{ex:schilderen}.\footnote{Of course, \tit{schilderen} also optionally takes an accusative DP, but I am focussing on the instrumental here.}

@@

\ref{ex:summaryconst} summarizes what I showed in this section. \tit{Met wat} `with what' can never surface when \tit{met} `with' and \tit{wat} `what' form a proper constituent. It always becomes \tit{waarmee} `with what'. This is schematically shown in \ref{ex:waarmeefr}.
There are other contexts in which \tit{met wat} `with what' appears. This can be either when \tit{wat} `what' takes a complement, or when \tit{wat} is part of the a clause that \tit{met} `with' is not a part of.This last option is schematically showed in \ref{ex:metwatfr}, and it represents the mismatching free relative in \ref{ex:mismatchmetwat}.

@@

\section{\tit{Waarmee} as suppletion}

@@

In what follows I show another instance in which proper constituency influences spellout, suppletive negative forms in Korean as discussed in \citet{chung2007}. The point is that suppletion only goes through when the relevant features form a proper constituent, and not when an additional feature intervenes.

@@

Spellout targets only constituent \citep{starke2009}. This ensures \ref{ex:entrymolu} is not inserted into the structure in \ref{ex:structurenegmolu}, as \tsc{caus} is contained in the syntactic structure but not in the structure in the lexicon in \ref{ex:entrymolu}.

@@

\section{A finer decomposition}

In this section I investigate the internal structure of \tit{waarmee} `with what' and \tit{met wat} `with what' to capture the phonological similarities between the two forms. First, I identify \tit{w-} and \tit{-a-} as morphemes that appear in both expressions. Putting these two aside, I concentrate on \tit{ermee} `with it' and \tit{met 't} `with it'. I decompose these two expressions further, and end up arguing for a picture as shown in Table \ref{tbl:nofeatures}.
The elements \tit{w-} and \tit{-a-} express the same syntactic structure in both expressions. The elements \tit{met} and \tit{'t} together also express the same features as \tit{mee} and \tit{er} together, but the distribution differs. \tit{Met} expresses less structure than \tit{mee} and \tit{'t} expresses more structure than \tit{er}. It becomes clear that the \tit{er} in fact corresponds to the locative \tit{r} in Dutch.

\begin{table}[ht]
	\center
	\caption {\tit{met wat} and \tit{waarmee}}
	\begin{minipage}{0.56\linewidth}
		\begin{tabularx}{\textwidth}{ccccccc}
		\toprule
    \phantom{\tsc{wh}}  & \phantom{\tsc{deix}}                    & \phantom{\tsc{f}4}  & \phantom{\tsc{f}3} & \phantom{\tsc{f}2}  & \phantom{\tsc{f}1}  & \phantom{\tsc{thing}} \\
		%\midrule
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \tit{met} & \multicolumn{4}{|c}{\tit{(ə)t}}                \\\hline
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \multicolumn{4}{c|}{\tit{mee}}               & \tit{(ə)r}  \\
    \bottomrule
\end{tabularx}
\end{minipage}
\end{table}\label{tbl:nofeatures}

@@

In this section I decompose \tit{waarmee} `with what' and \tit{met wat} `with what'. The idea is that both expessions realize the same features, and \tit{waarmee} `with what' takes precedence when all features form a constituent. The morphemes I distinguish in \tit{waarmee} `with what' are \tit{w-}, \tit{-aa-}, \tit{-r} and \tit{-mee}. Within \tit{met wat} `with what' I distinguish \tit{met}, \tit{w-}, \tit{-a-} and \tit{-t}. \ref{ex:decompose} shows this as well.

@@

Let me start with the morphemes \tit{w-} and \tit{-a-} that appear in both expressions. Exactly because they appear in both expression, I do not pay much attention to them. For \tit{w-} I follow \citet{hachem2015} in that \tit{w-} realizes a WP.

@@

The morpheme \tit{-a-} expresses deixis. Dutch distinguishes between proximal by using \tit{-ie-/-i-} and distal by using \tit{-aa-/-a-}, illustrated in \ref{ex:deixis}. I assume that lengthening or shortening of the vowel is a result of the final \tit{r}.

@@

In combination with wh-elements, the distal marker \tit{-a-} is used. For the purpose of this paper I simply let \tit{-a-} correspond to \tsc{deixP}.

@@

I put \tit{w-} and \tit{-a-} aside for now, assuming they spell out the same syntactic structure in \tit{waarmee} `with what' and \tit{met wat} `with what'. This leaves \tit{er-mee} `with it' and \tit{met 't} `with it'.\footnote{\tit{'rmee} `with it' is often precdeded by an /ε/, which I assume to be for phonological reasons. The vowel can be dropped under the right phonological circumstances.

\exg. Ik heb 'r niks mee te maken.\\
 I have there nothing with to do\\
 `I have nothing to do with it.'
 \z.

@@

\ex.
\ag. (e)r -mee\\
there with\\
\b. met 't\\
with it\\

@@

\ex. \begin{forest} boom
 [\tsc{dat}P
     [\tsc{f}3]
     [\tsc{acc}P
         [\tsc{f}2]
         [\tsc{nom}P
             [\tsc{f}1]
             [\tsc{thingP}
                 [\tsc{thing}, roof]
             ]
         ]
     ]
 ]
 {\draw (.east) node[right]{⇔ \tit{(ə)t}}; }
 \end{forest}\label{ex:entryt}

This lexical entry can lexicalize the \tsc{datP}, but also the \tsc{accP} and \tsc{nomP}. This is due to the Superset Princple.

@@

In other words, a lexically stored structure does not have to be identical to the syntactic structure. It is enough if the syntactic structure is contained in the lexically stored tree. This has a as consequence that the lexical entry in \ref{ex:entryt} can also be inserted in \ref{ex:tacc} and \ref{ex:tnom}.

@@

 ex. [ [ [ [ \tsc{thing} ] \tsc{f1} ] \tsc{f2} ] \tsc{f3} \tsc{f4} ]\label{ex:stmetwat}

@@

 \ex. \begin{forest} boom
 [\tsc{loc}P
 [\tsc{location}]
     [\tsc{thngP}
         [\tsc{thing}, roof]
     ]
 ]
 {\draw (.east) node[right]{⇔ \tit{(ə)r}}; }
 \end{forest}\label{ex:entryr}

Via the superset principle, \tit{er} can be used to realize the feature \tsc{thing} as well, as it is contained in \tsc{locP}.

So far, \tit{er} `there' only realizes the feature \tsc{thing} from the structure in \ref{ex:stmetwat}. This leaves \tit{f1} to \tsc{f4} to be realized by \tit{-mee}. I give the lexical tree in \ref{ex:entrymee}.

@@

Note that there are two differences between the lexical entry in \ref{ex:entrymet} and \ref{ex:entrymee}. First, the lexical entry for \tit{mee} `with' in \ref{ex:entrymee}, in addition to the one for \tit{met} `with' in \ref{ex:entrymet}, contains the features \tsc{f1}, \tsc{f2} and \tsc{f3}.
Second, the lexical entry in \ref{ex:entrymet} has a binary bottom, and the lexical entry in \ref{ex:entrymee} has a unary bottom. Nanosyntax distinguishes prefixes and suffixes by distinguishing these types of lexical entries. This is the crucial differences needed for the Dutch instrumental.
\tit{Met} `with' is a prefix (preposition), so it has a binary bottom, and \tit{-mee} `with' is a suffix, so it needs a unary bottom. In the next section I show how this works in a derivation.
\footnote{For completeness, \tit{mee} `with' has to be blocked from combining with full DPs and animate pronouns (reference). For now I assume that these pronouns spell out an additional feature \tit{σ} that is not realized by \tit{mee} `with' but only by \tit{met} `with'. Full DPs combine in addition with a zero marker for dative case. This way \tit{mee} `with' does not compete with \tit{met} `anymore'. I left this features out in this paper, at is does not contribute to the main point.}
\footnote{Giving \tit{met} `with' and \tit{mee} `with' two distinct lexical entries has as a consquence that the phonological overlap between them seems like a coincidence. I leave it for future research to capture this observation.}
\footnote{I do not address the ordering difference of the wh-element and the adposition. A topic related to (but not relevant for) this paper is the different positioning of identical adpositions in Dutch. In \ref{ex:dutchin}, \tit{in} changes meaning dependent on whether it procedes or follows the DP, it is respectively locational or directional.

@@

The different positioning of the adpositions in \tsc{r}-pronouns is driven by spellout and is, therefore, meaningless. However, the topic might be related. In almost all cases the preposition does not change form when it combines with an \tsc{r}-pronoun, e.g. \tit{in}. If this proposal is on the right track, such elements can be used as either a prefix and as a suffix. A lexical entry as in \ref{ex:presuf} would be a candidate for such an element.

I leave it to future research to determine whether this is a feasible solution.}

Table \ref{tbl:withfeatures} gives an overview of the distribution of the features in \tit{met wat} `with what' and \tit{waarmee} `with what'. In both cases, \tit{w-} and \tit{-a-} are the realization of the same morpheme. The difference between the two expressions lies in the distribution of the case features. \tit{Mee} `with' realizes all of them, and \tit{met} `with' only the top two. The residues are realized by respectively \tit{er} `there' and \tit{'t} `it'.

\begin{table}[ht]
	\center
	\caption {\tit{met wat} and \tit{waarmee}}
	\begin{minipage}{0.56\linewidth}
		\begin{tabularx}{\textwidth}{ccccccc}
		\toprule
    \tsc{wh}  & \tsc{deix}                     & \tsc{f}4  & \tsc{f}3 & \tsc{f}2  & \tsc{f}1  & \tsc{thing} \\
		\midrule
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \tit{met} & \multicolumn{4}{|c}{\tit{(ə)t}}                \\\hline
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \multicolumn{4}{c|}{\tit{mee}}               & \tit{(ə)r}  \\
    \bottomrule
\end{tabularx}
\end{minipage}
\end{table}\label{tbl:withfeatures}

In the next section I put all features back together in a derivation and I show how \tit{waarmee} `with what' surfaces when all features form a constituent. \tit{Met wat} `with what' appears when the functional sequence is disrubted.

>>>>

\title{\tit{Waarmee} vs. \tit{met wat}}
\author{Fenna Bergsma}
\date{\today}

@@

The term \tsc{r}-pronoun \citep{riemsdijk1978} refers to a set of nominal elements that can strand prepositions in Dutch (and German). This is noteworthy because Dutch is a normally non-preposition stranding language. \tsc{R}-pronouns morphologically consist of a combination of a preposition and locative pronoun. In the example \ref{ex:klimerop} \tit{op} `on' is the preposition and \tit{er} `there' is the locative pronoun.

@@

This paper focuses on a single \tsc{r}-pronoun in a specific type of free relative construction. The \tsc{r}-pronoun I focus on is the relative pronoun \tit{waarmee} `with what', which is interesting for two reasons. First, just like for all \tsc{r}-pronouns, the wh-element is the locative, but there is no meaning component related to location in `with what'. Second, the preposition \tit{met} `with' changes into \tit{mee} `with' when it is combined with an \tsc{r}-pronoun.

The construction I focus on is a mismatching free relative. This is a free relative construction in which the two predicates (the one in the main clause and the one in the embedded clause) combine with two different cases (i.e. the case requirements do not match). I illustrate this in \ref{ex:gekochtwaarmee}. The predicate in the embedded clause, \tit{schildert} `paint', combines with an instrumental PP. The predicate in the main clause clause, \tit{gekocht} `bought' combines with an accusative DP. The \tsc{r}-pronoun \tit{waarmee} `with what' is used here.\footnote{In this example \tit{mee} `with' is stranded and \tit{waar} `where' is moved to the left edge of the embedded clause. It is also possible for \tit{waar} `where' to bring \tit{mee} `with' along, as in \ref{ex:meealong}, but it is regarded as slightly less natural.

@@

If the predicates are switched around between the clauses, the \tsc{r}-pronoun does not appear anymore. In \ref{ex:schildermet}, \tit{schilder} `paint' combines with an instrumental PP in the main clause and \tit{gekocht} `bought' combines with an accusative DP in the embedded clause. The use of an \tsc{r}-pronoun is ungrammatical, as indicated by the ungrammaticality of \ref{ex:schilderwaarmee}. Instead, a combination of the regular instrumental preposition \tit{met} `with' and the regular wh-pronoun \tit{was} `what' in used.

@@

In this paper I show that distribution of \tit{waarmee} `with what' and \tit{met wat} `with what' in these free relative constructions gives us a good insight into the internal structure of \tsc{r}-pronouns. In what follows I show that \tsc{r}-pronouns and regular preposition compete to spell out the same syntactic features. If all features form a proper constituent (i.e. a constituent to the exclusion of other features), the \tsc{r}-pronoun surfaces. If it is not a proper constituent, the preposition-pronoun combination shows up. This straightforwardly follows in a system in which spellout targets phrasal constituents: Nanosyntax \citep{starke2009}.

This paper is structured as follows. In Section \ref{sec:distribution} I discuss the distribution of \tit{waarmee} `waarmee' and \tit{met wat} `with what'. I show that the \tsc{r}-pronoun surfaces when it forms a proper constituent, and \tit{met wat} `with what' when it does not. Section \ref{sec:suppletion} discusses how proper constituency relates to spellout by treating \tit{waarmee} `with what' as an instance of suppletion. In Section \ref{sec:finer}, I decompose \tit{waarmee} `with what' and \tit{met wat} `with what' to make sense of the phonological similarities, and I provide a more detailed analysis. Unmarked examples are constructed and have been verified by native speakers.

\section{The distribution between \tit{waarmee} and \tit{met wat}}\label{sec:distribution}

In the introduction I discussed the distribution between \tit{waarmee} `with what' and \tit{met wat} `with what' in free relatives with predicates that combine with different cases. Table \ref{tbl:distribution} repeats the generalization. When the main clause predicate combines with an accusative and the embedded clause predicate with an instrumental, \tit{waarmee} `with what' is grammatical and \tit{met wat} `with what' is ungrammatical. When the main clause predicate combines with an instrumental and the embedded clause predicate with an accusative, \tit{waarmee} `with what' is ungrammatical and \tit{met wat} `with what' is used.

\begin{table}[ht]
	\center
	\caption {Distribution between \tit{waarmee} and \tit{met wat}}
	\begin{minipage}{0.45\linewidth}
		\begin{tabularx}{\textwidth}{ccc}
		\toprule
                              & \tit{waarmee} & \tit{met wat} \\
		\midrule
    m:\tsc{acc}, e:\tsc{ins}  & ✔             & *             \\
    m:\tsc{ins}, e:\tsc{acc}  & *             & ✔             \\
    \bottomrule
\end{tabularx}
\label{tbl:distribution}
\end{minipage}
\end{table}

@@

 \ex. \label{ex:objverb}
 \ag. Ik zie haar/hem.\\
  I see her/him\\
  `I see her/him.'\label{ex:aniobj}
 \bg. Ik zie 't.\\
  I see it\\
  `I see it.'\label{ex:inaniobj}

@@

\ex. \label{ex:objprep}
\ag. Ik schilder samen met haar/hem.\\
 I paint together with her/him\\
 `I am painting together with her/him.'\label{ex:prepani}
\bg. *Ik schilder met 't.\\
 I paint with it\\
 `I am painting with it.'\label{ex:prephet}
\bg. Ik schilder er-mee.\\
 I paint there-with\\
 `I am painting with it.'\label{ex:preper}
\bg. *Ik schilder mee er.\\
 I paint with there\\
 `I am painting with it.'\label{ex:erprep}

@@

The situation of the inanimate wh-pronouns resembles the inanimate personal pronouns. \tit{Wat} `what' can function as an object of a verb (see \ref{ex:wat}), but not as an object of a preposition \ref{ex:metwat}. In that case, the \tsc{r}-pronoun \tit{waarmee} `with what' is appears, as shown in \ref{ex:waarmee}.\footnote{The sentence in \ref{ex:metwat} is unacceptable with neutral intonation. It becomes is only acceptable if \tit{wat} `what' is stressed, for example in a context in which the speaker is highly surprised about the choice for the object hearer is painting with.}

@@

\begin{table}[ht]
	\center
	\caption {Inanimates in Dutch}
	\begin{minipage}{0.7\linewidth}
		\begin{tabularx}{\textwidth}{ccc}
		\toprule
                        & pers. pronouns & wh-pronouns \\
  	\midrule
objects of verbs        & 't             & wat         \\
objects of prepositions & er             & waar        \\
\bottomrule
\end{tabularx}
\label{tbl:inanimates}
\end{minipage}
\end{table}

@@

In each of these examples the instrumental PP forms a constituent at a certain point in the derivation. In \ref{ex:const1}, the PP forms a proper constituent in the surface order, as shown in \ref{ex:const1stage}. In \ref{ex:const2}, the PP forms a proper constituent before wh- and V2- movement, shown in \ref{ex:const2stage}.
The structure in \ref{ex:const3stage} represents a stage in the derivation of the embedded clauses in \ref{ex:const3} and \ref{ex:const4}.\footnote{I assume that the antecedent in a free relative is a phonologically empty element, in line with cf. \cite{bresnan1978a,groos1981,himmelreich2017}. Under that view, \ref{ex:const3} and \ref{ex:const4} are identical, except for that in \ref{ex:const4} modifies phonologically empty element.}
Again, in the stage, which comes before relative movement of the pronoun to the left periphery of the relative clause, the PP forms a proper constituent.

@@

\ex.\label{ex:constwaarmee}
\ag. Waar-mee schilder jij?\\
 {where-with} paint you\\
 `What are you painting with?'
\bg. Ik schilder [waar-mee jij ook schildert].\\
 I paint {where-with} you also paint\\
 `I am painting with what you are painting with too.'
\bg. Ik schilder met de kwast [waar-mee jij ook schildert].\\
 I paint with the brush {where-with} you also paint\\
 `I am painting with the brush that you are painting with too.'

@@

Let me now show how this applies to the examples with the mismatching free relatives. The two predicates I used in the free relatives are \tit{kopen} `to buy' and \tit{schilderen}  `to paint'. \tit{Kopen} `to buy' takes an accusative DP as its object, illustrated in \ref{ex:kopen}. \tit{Schilderen} `to paint' can take an instrumental PP as its object., shown in \ref{ex:schilderen}.\footnote{\tit{Schilderen} also optionally takes an accusative object, but I am focussing on the instrumental here.}

@@

\ref{ex:summaryconst} summarizes what I showed in this section. \tit{Met wat} `with what' can never surface when \tit{met} `with' and \tit{wat} `what' form a proper constituent. It always becomes \tit{waarmee} `with what'. This is schematically shown in \ref{ex:waarmeefr}.
There are other contexts in which \tit{met wat} `with what' appears. This can be either when \tit{wat} `what' takes a complement, or when \tit{wat} is part of the a clause that \tit{met} `with' is not a part of. This last option is schematically showed in \ref{ex:metwatfr}, and it represents the mismatching free relative in \ref{ex:mismatchmetwat}.

@@

\section{\tit{Waarmee} `with what' as suppletion}\label{sec:suppletion}

@@

In what follows I show another instance in which proper constituency influences spellout, namely suppletive negative forms in Korean, as discussed by \citet{chung2007}. The point is that suppletion only goes through when the relevant features form a proper constituent, and not when an additional structure intervenes.

@@

Since spellout targets constituent \citep{starke2009}, \ref{ex:entrymolu} is not inserted into the structure in \ref{ex:structurenegmolu}. \tsc{caus} is contained in the syntactic structure but not in the structure in the lexicon in \ref{ex:entrymolu}.

@@

\section{A finer decomposition}\label{sec:finer}

In this section I investigate the internal structure of \tit{waarmee} `with what' and \tit{met wat} `with what' to capture the phonological similarities between the two forms. First, I identify \tit{w} and \tit{a} as morphemes that appear in both expressions. Putting these two aside, I concentrate on \tit{ermee} `with it' and \tit{met 't} `with it'. I decompose these two expressions further, and end up arguing for a picture as shown in Table \ref{tbl:nofeatures}.
The elements \tit{w} and \tit{a} express the same syntactic structure in both expressions. The elements \tit{met} and \tit{'t} together also express the same features as \tit{mee} and \tit{er} together, but the distribution differs. \tit{Met} expresses less structure than \tit{mee} and \tit{'t} expresses more structure than \tit{er}. It becomes clear that the \tit{er} in fact corresponds to the locative \tit{r} in Dutch.

\begin{table}[ht]
	\center
	\caption {\tit{met wat} and \tit{waarmee}}
	\begin{minipage}{0.56\linewidth}
		\begin{tabularx}{\textwidth}{ccccccc}
		\toprule
    \phantom{\tsc{wh}}  & \phantom{\tsc{deix}}                    & \phantom{\tsc{f}4}  & \phantom{\tsc{f}3} & \phantom{\tsc{f}2}  & \phantom{\tsc{f}1}  & \phantom{\tsc{thing}} \\
		%\midrule
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \tit{met} & \multicolumn{4}{|c}{\tit{'t}}                \\\hline
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \multicolumn{4}{c|}{\tit{mee}}               & \tit{er}  \\
    \bottomrule
\end{tabularx}
\label{tbl:nofeatures}
\end{minipage}
\end{table}

@@

In this section I decompose \tit{waarmee} `with what' and \tit{met wat} `with what'. The idea is that both expressions realize the same features, and \tit{waarmee} `with what' takes precedence when all features form a constituent. The morphemes I distinguish in \tit{waarmee} `with what' are \tit{w}, \tit{-aa-}, \tit{-r} and \tit{mee}. Within \tit{met wat} `with what' I distinguish \tit{met}, \tit{w}, \tit{a} and \tit{-t}. \ref{ex:decompose} shows this as well.

@@

Let me start with the morphemes \tit{w} and \tit{a} that appear in both expressions. Exactly because they appear in both expression, I do not pay much attention to them. For \tit{w} I follow \citet{hachem2015} in that \tit{w} realizes a WP.

@@

The morpheme \tit{a} expresses deixis. Dutch distinguishes between proximal by using \tit{ie/i} and distal by using \tit{aa/a}, illustrated in \ref{ex:deixis}. I assume that lengthening or shortening of the vowel is a result of the final \tit{r}.

@@

In combination with wh-elements, the distal marker \tit{a} is used. For the purpose of this paper I simply let \tit{a} correspond to \tsc{deixP}.

@@

I put \tit{w} and \tit{a} aside for now, assuming they spell out the same syntactic structure in \tit{waarmee} `with what' and \tit{met wat} `with what'. This leaves \tit{er-mee} `with it' and \tit{met 't} `with it'.\footnote{I assume that the \tit{e} in \tit{ermee} `with it' is there for phonological reasons. The vowel is dropped under the right phonological circumstances.

\exg. Ik he niks mee te maken.\\
 I have there nothing with to make\\
 `I have nothing to do with it.'
 \z.

@@

\ex.
\ag. er -mee\\
there with\\
\b. met 't\\
with it\\

@@

\ex. \begin{forest} boom
 [\tsc{dat}P
     [\tsc{f}3]
     [\tsc{acc}P
         [\tsc{f}2]
         [\tsc{nom}P
             [\tsc{f}1]
             [\tsc{thingP}
                 [\tsc{thing}, roof]
             ]
         ]
     ]
 ]
 {\draw (.east) node[right]{⇔ \tit{'t}}; }
 \end{forest}\label{ex:entryt}

This lexical entry can lexicalize the \tsc{datP}, but also the \tsc{accP} and \tsc{nomP}. This is due to the Superset Principle.

@@

In other words, a lexically stored structure does not have to be identical to the syntactic structure. It is enough if the syntactic structure is contained in the lexically stored tree. This has a as consequence that the lexical entry in \ref{ex:entryt} can also be inserted in syntactic structures as \ref{ex:tacc} and \ref{ex:tnom}.

@@

\ex. [ [ [ [ \tsc{thing} ] \tsc{f1} ] \tsc{f2} ] \tsc{f3} \tsc{f4} ]\label{ex:stmetwat}

@@

 \ex. \begin{forest} boom
 [\tsc{loc}P
 [\tsc{location}]
     [\tsc{thngP}
         [\tsc{thing}, roof]
     ]
 ]
 {\draw (.east) node[right]{⇔ \tit{er}}; }
 \end{forest}\label{ex:entryr}

Notice already here that, via the superset principle, \tit{er} can be used to realize the feature \tsc{thing} as well, as it is contained in \tsc{locP}.

So far, \tit{er} `there' only realizes the feature \tsc{thing} from the structure in \ref{ex:stmetwat}. This leaves \tsc{f1} to \tsc{f4} to be realized by \tit{mee}. I give the lexical tree in \ref{ex:entrymee}.

@@

Note that there are two differences between the lexical entry in \ref{ex:entrymet} and \ref{ex:entrymee}. First, the lexical entry for \tit{mee} `with' in \ref{ex:entrymee}, in addition to the one for \tit{met} `with' in \ref{ex:entrymet}, contains the features \tsc{f1}, \tsc{f2} and \tsc{f3}.
Second, the lexical entry in \ref{ex:entrymet} has a binary bottom, and the lexical entry in \ref{ex:entrymee} has a unary bottom. Nanosyntax distinguishes prefixes and suffixes by distinguishing these types of lexical entries. This is the crucial differences needed for the Dutch instrumental.
\tit{Met} `with' is a prefix (preposition), so it has a binary bottom, and \tit{mee} `with' is a suffix, so it needs a unary bottom. In the next section I show how this works in a derivation.\footnote{For completeness, \tit{mee} `with' has to be blocked from combining with full DPs and animate pronouns (reference). For now I assume that these pronouns spell out an additional feature \tit{σ} that is not realized by \tit{mee} `with' but only by \tit{met} `with'. Full DPs combine in addition with a zero marker for dative case. This way \tit{mee} `with' does not compete with \tit{met} `anymore'. I left this features out in this paper, at is does not contribute to the main point.}\footnote{Giving \tit{met} `with' and \tit{mee} `with' two distinct lexical entries has as a consequence that the phonological overlap between them seems like a coincidence. Moreover, in almost all cases the preposition does not change form when it combines with an \tsc{r}-pronoun, e.g. \tit{in}. If this proposal is on the right track, elements as \tit{in} can be used as either a prefix and as a suffix. A lexical entry as in \ref{ex:presuf} would be a candidate for such an element.

I leave it to future research to determine whether this is a feasible solution.}\footnote{I do not address the ordering difference of the wh-element and the adposition. A topic related to (but not relevant for) this paper is the different positioning of identical adpositions in Dutch. In \ref{ex:dutchin}, \tit{in} changes meaning dependent on whether it proceeds or follows the DP, it is respectively locational or directional.

@@

The different positioning of the adpositions in \tsc{r}-pronouns is driven by spellout and is, therefore, meaningless.}

Table \ref{tbl:withfeatures} gives an overview of the distribution of the features in \tit{met wat} `with what' and \tit{waarmee} `with what'. In both cases, \tit{w} and \tit{a} are the realization of the same morpheme. The difference between the two expressions lies in the distribution of the case features. \tit{Mee} `with' realizes all of them, and \tit{met} `with' only the top two. The residues are realized by respectively \tit{er} `there' and \tit{'t} `it'.

\begin{table}[ht]
	\center
	\caption {\tit{met wat} and \tit{waarmee}}
	\begin{minipage}{0.56\linewidth}
		\begin{tabularx}{\textwidth}{ccccccc}
		\toprule
    \tsc{wh}  & \tsc{deix}                     & \tsc{f}4  & \tsc{f}3 & \tsc{f}2  & \tsc{f}1  & \tsc{thing} \\
		\midrule
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \tit{met} & \multicolumn{4}{|c}{\tit{'t}}                \\\hline
    \tit{w}   & \multicolumn{1}{|c|}{\tit{a}}  & \multicolumn{4}{c|}{\tit{mee}}               & \tit{er}  \\
    \bottomrule
\end{tabularx}
\label{tbl:withfeatures}
\end{minipage}
\end{table}

In the next section I put all features back together in a derivation and I show how \tit{waarmee} `with what' surfaces when all features form a constituent. \tit{Met wat} `with what' appears when the functional sequence is disrupted.
